Deciphering the Relationship Betweeen Race, Income, and Political Support in the 2020 United States Election

### 2024-03-15 paper/paper.qmd
model section nearly done

The CES 2020 dataset contained a sample size of approximately 61,000 cases. This extensive collection reflects a broad spectrum of variables, covering political, demographic, and socio-economic data points. Nevertheless, it was noted that certain variables had missing values or were not directly related to my examination. In the initial stages of data cleaning, attention was concentrated on crucial variables such as voter identification, voter registration status, race, and income. Variables that closely corresponded to this paper, including those related to demographics and electoral participation, were prioritized. Whereas, variables with less direct relevance to the paper were omitted to efficiently analyze the dataset.

@@

# Model

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

@@

### Model justification

>>>>

The CES 2020 dataset contained a sample size of approximately 61,000 cases. This extensive collection reflects a broad spectrum of variables, covering political, demographic, and socio-economic data points. Nevertheless, it was noted that certain variables had missing values or were not directly related to my examination. In the initial stages of data cleaning, attention was concentrated on crucial variables such as voter identification, voter registration status, presidential preferences, voter turnout, race, and income. Variables that closely corresponded to this paper, including those related to demographics and electoral participation, were prioritized. Whereas, variables with less direct relevance to the paper were omitted to efficiently analyze the dataset.

@@

# Model
To investigate the factors influencing voting behavior within the Cooperative Election Study (CES) 2020 dataset, we employed a logistic regression model, leveraging the `rstanarm` package of @rstanarm for its Bayesian inference capabilities. This choice allows for a more nuanced interpretation of results, incorporating prior beliefs and providing probabilistic estimates of the parameters. Secondly, this model aims to unpack the impact of demographic variables such as race and income on the probability of an individual casting a vote, framed within a binary outcome (voted vs. did not vote). The overarching goal of our modeling strategy is to distill the complex interplay of these demographic variables, helping our understanding of voting behaviors.

@@

Define $y_i$ as the political preference of the respondent and equal to 1 if Biden and 0 if Trump. Then $\mbox{race}_i$ is the race of the respondent and $\mbox{income}_i$ is the income of the respondent.

\begin{aligned}
y_i|\pi_i & \sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) & = \beta_0+\beta_1 \times \mbox{race}_i + \beta_2 \times \mbox{income}_i \\
\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
\beta_2 & \sim \mbox{Normal}(0, 2.5)
\end{aligned}

@@

### Model justification
For our model, we expect more minorities to vote for Biden based on his campaign. We also expect there to be a considerable effect of income. We intend to investigate how these demographic factors may influence the probability of individuals voting for a particular candidate. To convey these relationships mathematically, we can represent them using LaTeX within dollar signs, such as $\beta_1$ for the effect of race and $\beta_2$ for the effect of income.
